### 2025-02-28 Manuscript_Drafts.Rmd
manuscript stuff

---
title: "manuscript_stuff"
author: "Kristen Johnson"
date: "2025-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

OPTION 1:

Introduction: The Feathered Balance Sheet: A Mixed Methods Examination of Urban Backyard Chicken Husbandry
In recent decades, the practice of maintaining domestic poultry within urban and suburban residential settings has experienced a notable resurgence. This phenomenon, colloquially referred to as "backyard chicken keeping," represents a fascinating intersection of agricultural tradition and contemporary sustainability movements. The economic landscape has further accelerated this trend, as inflated egg prices have prompted an increasing number of households to consider domestic poultry as a pragmatic response to marketplace volatility. Indeed, consumer data suggests that the correlation between rising commercial egg costs and the purchase of chicken coops is remarkably robust.
The quantitative and qualitative dimensions of this investigation address a critical gap in the literature, which has hitherto focused either on large-scale commercial poultry operations or on historical agricultural practices. While proponents extol the virtues of fresh eggs and animated garden assistants, detractors cite concerns ranging from zoonotic disease transmission to neighborhood noise pollution—a concern that anyone who has experienced a rooster's enthusiasm for announcing dawn can readily appreciate.
Current Study
Our research design employs a mixed methods approach to develop a nuanced understanding of this increasingly common practice. The methodology combines quantitative data analysis examining multiple factors related to backyard chicken husbandry—including cost-benefit analyses, time investment, and zoning compliance—with qualitative thematic analysis of in-depth interviews conducted with chicken owners across three metropolitan areas. By triangulating these methodological approaches, we aspire to move beyond anecdotal evidence and develop an empirically robust understanding of what might be termed the "net hennefit" of backyard chicken husbandry in contemporary urban contexts. The present study thus investigates the hypothesis that the benefits of maintaining a small flock of Gallus gallus domesticus in residential settings significantly outweigh the associated challenges and drawbacks.

OPTION 2:

Introduction
Backyard chicken keeping has surged in popularity in recent years, driven by a desire for sustainable food production, self-sufficiency, and ethical animal husbandry (_______CITE_________). While much of the existing literature focuses on small and medium-sized breeds, large chicken breeds—such as Brahmas, Jersey Giants, and Orpingtons—offer unique advantages and challenges for backyard poultry enthusiasts. These birds often provide higher egg production, greater meat yield, and a more docile temperament, making them attractive to small-scale farmers and homesteaders (_______CITE_________). However, their size also necessitates specialized care, including adjustments in nutrition, housing, and health management.  
Despite the growing interest in raising large chickens in non-commercial settings, research remains limited on how backyard keepers manage the specific needs of these birds. Understanding the factors that contribute to their well-being—such as space requirements, dietary considerations, and health monitoring—is crucial for promoting sustainable and humane poultry practices. Furthermore, the motivations, experiences, and challenges of backyard chicken keepers remain underexplored, particularly regarding the balance between productivity and ethical care (_______CITE_________).  
This study employs a mixed methods approach to examine the care and management of large chicken breeds in backyard settings. Using surveys and interviews with backyard poultry keepers, alongside observational data on bird health and environmental conditions, we aim to identify best practices and common challenges in raising these birds. By integrating qualitative and quantitative data, this research seeks to provide a comprehensive understanding of how backyard poultry keepers navigate the complexities of raising large chickens. The findings will contribute to the broader discourse on sustainable animal husbandry and inform best practices for individuals seeking to maintain healthy, productive backyard flocks.