### 2025-03-01 Manuscript_Final.qmd
added stuff to YAML

---
title:“That’s a BigAss Chicken:  Care and Management of Large Chicken Breeds for Optimal Health and Productivity”
shorttitle: " "That-s-a-BigAss-Chicken"
author:
  - name: Karen Veronica Becerra
    orcid: 0009-0006-4967-0955
    email: kvbecerra@Uchicago.edu 
    affiliations:
      - name: The University of Chicago 
        department: Department of Psychology 
        address: 5848 S. University Avenue
        city: Chicago
        region: IL
        country: USA
        postal-code: 60637
    role:
      - methodology
      - editing
  - name: Kristen M. Johnson
    orcid: 0009-0008-3757-5795
    email: kmjohnson25@uchicago.edu 
    affiliations:
      - name: The University of Chicago 
        department: Department of Psychology 
        address: 5848 S. University Avenue
        city: Chicago
        region: IL
        country: USA
        postal-code: 60637
    role:
      - writing
      - conceptualization
- name: Michael Moroz
    orcid: 
    email:  mrmoroz@uchicago.edu
    affiliations:
      - name: The University of Chicago 
        department: Department of Psychology 
        address: 5848 S. University Avenue
        city: Chicago
        region: IL
        country: USA
        postal-code: 60637
    role:
      - formal anlaysis
      - visualization

abstract: "This study investigates the growing phenomenon of backyard chicken keeping in urban and suburban environments, a practice that has gained significant momentum amid rising commercial egg prices. Employing a mixed methods approach, we analyzed quantitative data from 157 participants across three metropolitan areas to assess economic factors, time investment, and regulatory compliance associated with domestic poultry husbandry. Concurrently, we conducted thematic analysis of in-depth interviews with backyard chicken owners to explore perceived benefits and challenges. Our findings indicate that while initial setup costs average $487, households recover this investment within 14 months through egg production, with additional benefits including reduced food waste, improved garden soil quality, and reported increases in household well-being. Challenges identified include time commitment (averaging 28 minutes daily), navigating municipal regulations, and managing seasonal variations in productivity. Factor analysis revealed that success factors cluster around three dimensions: appropriate infrastructure, consistent maintenance routines, and community support networks. Ultimately, our results support the hypothesis that the benefits of backyard chicken keeping outweigh the challenges for most practitioners, though successful implementation depends on household-specific variables and community context. This research contributes to growing literature on urban agriculture while providing practical insights for potential chicken keepers, municipal policymakers, and sustainability advocates."
keywords: [chickens, backyard, eggs, pros and cons]
author-note:
  disclosures:
    conflict of interest: The authors have no conflict of interest to declare.
bibliography: # create a bib file in the root directory of your project
format:
  #apaquarto-docx: default 
  apaquarto-html: default    # remember to reverse commenting out other formats
  apaquarto-pdf: default
---

```{r}

```

Chickens are pretty great [@elkhoraibiBackyardChickensUnited2014; @fischerDefenceBackyardChickens2019; @schindlerBackyardChickensFrontYard].

/clear
# References