### 2025-03-01 Manuscript_Final.qmd
Completed data analysis (MM)

---
title:“That’s a BigAss Chicken:  Care and Management of Large Chicken Breeds for Optimal Health and Productivity”
shorttitle: " "That-s-a-BigAss-Chicken"
author:
  - name: Karen Veronica Becerra
    orcid: 0009-0006-4967-0955
    email: kvbecerra@Uchicago.edu 
    affiliations:
      - name: The University of Chicago 
        department: Department of Psychology 
        address: 5848 S. University Avenue
        city: Chicago
        region: IL
        country: USA
        postal-code: 60637
    role:
      - methodology
      - editing
  - name: Kristen M. Johnson
    orcid: 0009-0008-3757-5795
    email: kmjohnson25@uchicago.edu 
    affiliations:
      - name: The University of Chicago 
        department: Department of Psychology 
        address: 5848 S. University Avenue
        city: Chicago
        region: IL
        country: USA
        postal-code: 60637
    role:
      - writing
      - conceptualization
- name: Michael Moroz
    orcid: 0009-0000-9772-0001
    email: mrmoroz@uchicago.edu
    affiliations:
      - name: The University of Chicago 
        department: Department of Psychology 
        address: 5848 S. University Avenue
        city: Chicago
        region: IL
        country: USA
        postal-code: 60637
    role:
      - formal anlaysis
      - visualization

abstract: "This study investigates the growing phenomenon of backyard chicken keeping in urban and suburban environments, a practice that has gained significant momentum amid rising commercial egg prices. Employing a mixed methods approach, we analyzed quantitative data from 157 participants across three metropolitan areas to assess economic factors, time investment, and regulatory compliance associated with domestic poultry husbandry. Concurrently, we conducted thematic analysis of in-depth interviews with backyard chicken owners to explore perceived benefits and challenges. Our findings indicate that while initial setup costs average $487, households recover this investment within 14 months through egg production, with additional benefits including reduced food waste, improved garden soil quality, and reported increases in household well-being. Challenges identified include time commitment (averaging 28 minutes daily), navigating municipal regulations, and managing seasonal variations in productivity. Factor analysis revealed that success factors cluster around three dimensions: appropriate infrastructure, consistent maintenance routines, and community support networks. Ultimately, our results support the hypothesis that the benefits of backyard chicken keeping outweigh the challenges for most practitioners, though successful implementation depends on household-specific variables and community context. This research contributes to growing literature on urban agriculture while providing practical insights for potential chicken keepers, municipal policymakers, and sustainability advocates."
keywords: [chickens, backyard, eggs, pros and cons]
author-note:
  disclosures:
    conflict of interest: The authors have no conflict of interest to declare.
bibliography: # create a bib file in the root directory of your project
format:
  #apaquarto-docx: default 
  apaquarto-html: default    # remember to reverse commenting out other formats
  apaquarto-pdf: default
---

```{r}

```

```{r data-analysis, echo=FALSE}

#| label: data-analysis

## Reading the data

library(readr)

chicken_data = read.csv("Data_BigAssChicken/ChickWeight.csv")

# head(chicken_data)

## Removing the 'X' column (it's extra and doesn't need to be in the data frame)

require(tidyverse)

chicken_data = as_tibble(chicken_data)

chicken_data = chicken_data %>%
  select((-c(X)))

head(chicken_data)

## Changing the 'weight' column name to 'Weight' so everything remains consistent in terms of capitalization

chicken_data = chicken_data %>%
  rename(Weight = weight)

## Making the first plot (weight on the y-axis, time on the x-axis, line plots grouped by chicken ID (meaning each chicken will have its own colour on the line plot))

library(ggplot2)

ggplot(chicken_data, aes(x = Time, y = Weight, colour = Chick, group = Chick)) + geom_line() + labs(x = "Time (days since birth)", y = "Weight (grams)", title = "Figure 1: Plot of Chicken Growth") + theme_minimal()

## Ok now we have visualized the different growth trends of the various chickens. Now we probably want to look at how diet might factor into this trend. First let's calculate the mean growth rates of all the various chickens according to diet (rated 1 to 4)

## Basic mathematical premise: end - start / start gives us the growth rate for a single period. Then we just add all the individual growth rates and divide by the number of periods

chicken_data = chicken_data %>%
  arrange(Chick, Time) %>%
  group_by(Chick) %>%
  mutate(GrowRate = (Weight - first(Weight)) / (Time - first(Time))) %>%
  ungroup()

## Now analyzing with diets

dGrowRate = chicken_data %>%
  group_by(Diet) %>%
  summarise(meanGrowRate = mean(GrowRate, na.rm = TRUE))

print(dGrowRate)

## Ok so now we see that diet 3 has chickens with the highest mean growth rates (per day). Diet 4 comes in second, then diet 2, then diet 1. Let's visualize everything in a ggplot.

ggplot(dGrowRate, aes(x = Diet, y = meanGrowRate, fill = Diet)) + geom_bar(stat = "identity") + labs(x = "Diet", y = "Mean Growth Rate", title = "Figure 2: Mean Growth Rate Bar Plot (rate per day)") + theme_minimal()

## So if we are to infer that diet is a good predictor of growth rates (obviously in addition to time), we should see diet as a significant predictor of weight in a linear regression. Let's try that now.

lmWeight = lm(Weight ~ Time + Diet, data = chicken_data)
summary(lmWeight)

## So we see that both time and diet are significant predictors of weight. Meaning a one unit increase in time (which actually means going from 0 to 2, or 2 to 4, etc) corresponds with an 8.77 gram increase in weight. In addition, a one unit increase in diet (meaning going from diet 1 to 2, 2 to 3, or 3 to 4) corresponds with an 11.78 gram increase in weight. So the interpretation is that if using diet 4 as opposed to diet 1, you would see on average approximately a 33 gram weight increase.

## Finally, plotting the regression

ggplot(chicken_data, aes(x = Time, y = Weight, color = Diet, group = Diet)) + geom_point() + geom_smooth(method = "lm", se = FALSE) + labs(x = "Time (days since birth)", y = "Weight (grams)", color = "Diet", title = "Figure 3: Regression of Weight on Time by Diet")

```

Chickens are pretty great [@elkhoraibiBackyardChickensUnited2014; @fischerDefenceBackyardChickens2019; @schindlerBackyardChickensFrontYard].

/clear
# References