### 2025-03-04 BAC_Qualitative-Data.Rmd
added all the qualitative stuff (maybe went overboard in imagining owning chickens)

---
title: "BAC_Qualitative-Data"
author: "Kristen Johnson"
date: "2025-03-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [MADE UP] QUALITATIVE DATA: Thematic Analysis of Interviews with Urban Backyard Chicken Farmers

## Methodology

Semi-structured interviews were conducted with 50 urban backyard chicken keepers (32 female, 17 male, 1 non-binary) across three metropolitan areas (Eastern Seaboard, n=20; Midwest, n=15; Pacific Northwest, n=15). Participants had maintained backyard chickens for periods ranging from 6 months to 12 years (mean = 3.7 years). Flock sizes ranged from 2-15 birds (median = 4). Interviews lasting 45-90 minutes were audio-recorded, transcribed verbatim, and analyzed using Braun and Clarke's (2006) six-phase approach to thematic analysis. Initial coding was conducted independently by two researchers, followed by collaborative theme development. Member-checking was employed with a subset of 12 participants to verify thematic authenticity.

## Themes and Subthemes

### 1. Economic Pragmatism

#### 1.1 Response to Market Volatility
Participants frequently cited economic considerations as initial motivations for chicken keeping, with 38 participants (76%) specifically mentioning rising egg prices:

> "When eggs hit seven dollars a dozen last winter, that was my tipping point. I did the math and realized a coop would pay for itself within a year." (Participant 23, Midwest)

#### 1.2 Beyond Simple Cost Calculations
While economic benefits served as initial motivations, participants described a more complex value proposition that extended beyond mere cost savings:

> "It's not just about saving money on eggs anymore. There's value in knowing exactly what my birds are eating, how they're treated, and connecting with my food source." (Participant 7, Eastern Seaboard)

#### 1.3 Unexpected Economic Benefits
Many participants (n=29, 58%) reported unanticipated economic advantages, including reduced food waste, garden fertilization, and pest control:

> "They eat all our kitchen scraps, turn them into eggs and fantastic compost. My garden has never been more productive, which means even more grocery savings." (Participant 42, Pacific Northwest)

### 2. Self-Sufficiency and Food Security

#### 2.1 Control Over Food Supply
A dominant theme emerged around increased feelings of food security and self-reliance:

> "After supply chain disruptions during COVID, having my own egg supply feels like insurance. No matter what happens at the grocery store, we've got breakfast." (Participant 11, Eastern Seaboard)

#### 2.2 Quality and Freshness
Nearly all participants (n=47, 94%) emphasized superior egg quality as a significant benefit:

> "Store-bought eggs are a completely different food. Once you've had eggs with those vibrant orange yolks from your backyard, there's no comparison." (Participant 39, Pacific Northwest)

#### 2.3 Stepping Stone to Greater Self-Reliance
Many participants (n=31, 62%) described chicken keeping as part of a broader trajectory toward increased self-sufficiency:

> "The chickens were our gateway. Now we've got vegetable beds, rainwater collection, and we're thinking about beekeeping next year." (Participant 28, Midwest)

### 3. Social and Community Dimensions

#### 3.1 Neighborhood Connections
Unexpectedly, chicken keeping frequently facilitated community building:

> "Our chickens are neighborhood celebrities. Kids stop by after school, neighbors trade garden produce for eggs, and we've met people on our street we never spoke to before." (Participant 3, Eastern Seaboard)

#### 3.2 Educational Value
Families with children (n=22) universally cited educational benefits:

> "My kids understand where food comes from. They've witnessed the whole life cycle, learned responsibility through daily care, and developed empathy for animals." (Participant 45, Pacific Northwest)

#### 3.3 Negotiating Regulations and Relationships
Navigating municipal regulations and neighbor relationships emerged as significant challenges:

> "Getting the permit was straightforward, but keeping peace with one particular neighbor has been ongoing diplomacy. Giving them free eggs helps." (Participant 19, Midwest)

### 4. Practical Challenges and Adaptations

#### 4.1 Time Commitment Reality
Time requirements emerged as the most frequently cited challenge (n=41, 82%):

> "The daily care is minimal, maybe 10 minutes. But when something goes wrong—a predator gets in, or a bird gets sick—suddenly it's very time-intensive." (Participant 34, Pacific Northwest)

#### 4.2 Seasonal Variability
Participants described adapting to natural laying cycles:

> "Winter egg production drops dramatically. You either accept buying eggs for a few months or install coop lighting, which raises ethical questions about manipulating their natural cycles." (Participant 8, Eastern Seaboard)

#### 4.3 Learning Curve and Resource Networks
Most participants (n=43, 86%) described a significant learning curve, mitigated by online communities:

> "YouTube and Facebook groups were lifesavers. There's always someone online who's dealt with whatever weird chicken situation you're facing at 11 PM." (Participant 14, Midwest)

### 5. Psychological and Emotional Benefits

#### 5.1 Stress Reduction
An unexpected theme emerged around the stress-reducing properties of chicken keeping:

> "Watching the chickens scratch and peck is meditative. After a stressful workday, I'll sit with a cup of tea and just observe them. Better than therapy." (Participant 31, Midwest)

#### 5.2 Connection to Agricultural Heritage
Many participants (n=26, 52%) described chicken keeping as connecting them to family heritage:

> "My grandmother kept chickens. When I collect eggs, I feel connected to her and generations of women in my family who did this same simple act." (Participant 5, Eastern Seaboard)

#### 5.3 Identity and Values Alignment
Chicken keeping frequently represented an expression of environmental and ethical values:

> "It's living my values around humane animal treatment, reducing my carbon footprint, and stepping outside the industrial food system, even if in a small way." (Participant 47, Pacific Northwest)

## Intersection of Themes

The interrelationship between themes reveals that while economic considerations often initiated chicken keeping, sustained practice was motivated by a complex interplay of factors. Particularly notable was the transformation from primarily economic motivation to deeper psychological, social, and philosophical benefits over time. As one participant summarized:

> "We started for the eggs, but we continue for everything else—the compost, the joy they bring to our kids, the neighbors they've introduced us to, and the simple pleasure of caring for animals that provide for us in return." (Participant 17, Midwest)

## Conclusion

This thematic analysis reveals backyard chicken keeping as a multidimensional practice extending well beyond utilitarian egg production. The initial economic calculus frequently evolves into a complex set of values-based, social, and psychological benefits that practitioners find equally or more compelling than the original financial motivation. Challenges are acknowledged but generally framed as manageable through adaptation, learning, and community support. These findings support our quantitative results indicating that practitioners perceive the benefits of backyard chicken keeping to outweigh the associated challenges, though with important nuance regarding the evolving nature of what constitutes "benefit" over the course of the practice.