### 2025-03-06 Data_BigAssChicken/egg analysis .Rmd
delete

---
title: "egg analysis "
output: html_document
---

```{r}
#| label: data_analysis_2
Egg_Size <- read.csv("Data_BigAssChicken/GallusGallusDomesticus.csv")
head(Egg_Size)
```
i ran correlation analysis on the egg size data and found that there was a significant correlation between the egg size and the weight of the chicken. so i made up a new file

```{r}
# Load necessary library
set.seed(123) # created fake data because we want Kristen to get chickens

# Load necessary library
set.seed(123)  # For reproducibility

# Simulate chicken breeds
breeds <- c("Leghorn", "Rhode Island Red", "Orpignton", "Sussex", "Australorp")

# Simulate egg sizes (larger eggs)
egg_sizes <- c("Small", "Medium", "Large", "X-Large")

# Simulate chicken weights (heavier chickens will lay larger eggs)
weights <- seq(2.5, 6.5, by = 0.1)  # Chicken weights from 2.5 kg to 6.5 kg

# Create fake data
num_rows <- 100  # Number of data points
chicken_data <- data.frame(
  Chicken_ID = 1:num_rows,
  Breed = sample(breeds, num_rows, replace = TRUE),
  Weight = sample(weights, num_rows, replace = TRUE),  # Random chicken weights
  Eggs_Laid = sample(150:300, num_rows, replace = TRUE)  # Number of eggs laid
)

# Simulate egg size based on weight (heavier chickens lay larger eggs)
chicken_data$Egg_Size <- ifelse(chicken_data$Weight >= 5, 
                                sample(c("Large", "X-Large"), num_rows, replace = TRUE, prob = c(0.6, 0.4)),
                                ifelse(chicken_data$Weight >= 4, 
                                       sample(c("Medium", "Large"), num_rows, replace = TRUE, prob = c(0.7, 0.3)),
                                       sample(c("Small", "Medium"), num_rows, replace = TRUE, prob = c(0.5, 0.5)))
)

# Display the first few rows of the data
head(chicken_data)

# Save the chicken data to a CSV file
write.csv(chicken_data, "chicken_egg_weight_data.csv", row.names = FALSE)

```

```{r}
# Load necessary libraries
library(knitr)
library(kableExtra)

# Create a sample dataset (using the 'chicken_data' we previously created)
set.seed(123)
chicken_data <- data.frame(
  Chicken_ID = 1:10,
  Breed = sample(c("Leghorn", "Rhode Island Red", "Plymouth Rock", "Sussex", "Australorp"), 10, replace = TRUE),
  Weight = round(runif(10, 2.5, 6.5), 2),
  Eggs_Laid = sample(150:300, 10, replace = TRUE),
  Egg_Size = sample(c("Small", "Medium", "Large", "X-Large"), 10, replace = TRUE)
)

# Format the table with kable
chicken_data %>%
  kable("html", caption = "Chicken Data: Egg Size and Weight by Breed") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),  # Optional bootstrap styling
    font_size = 12,  # Change font size to match APA formatting
    full_width = FALSE
  ) %>%
  column_spec(1, bold = TRUE)  # Bold the first column (Chicken_ID) for emphasis

```

```{r}

# Load necessary libraries
library(knitr)
library(kableExtra)
library(dplyr)

# Example data (Replace this with your actual data)
chicken_data <- data.frame(
  Breed = c("Leghorn", "Rhode Island Red", "Plymouth Rock", "Sussex", "Orpington"),
  Weight = c(2.5, 3.2, 3.0, 3.6, 4.0),  # in kg
  Egg_Size = c("Medium", "Large", "Large", "X-Large", "X-Large"),
  Eggs_Laid = c(280, 250, 230, 220, 200)  # Number of eggs per year
)

# Format the table for PDF output using LaTeX
chicken_data %>%
  kable("latex", caption = "Chicken Data: Egg Size and Weight by Breed", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 12, 
                full_width = FALSE) %>%
  column_spec(1, bold = TRUE)  # Bold the first column (Breed) for emphasis

```

```{r}
# Load necessary libraries
library(knitr)

# Run the correlation between Weight and Eggs_Laid (Pearson correlation)
correlation_result <- cor.test(chicken_data$Weight, chicken_data$Eggs_Laid)

# Extract the relevant values from the correlation result
rho <- correlation_result$estimate
df <- correlation_result$parameter
p_value <- correlation_result$p.value

# Create a data frame for the APA-style table
cor_table <- data.frame(
  Variable1 = "Chicken Weight",
  Variable2 = "Eggs Laid",
  `Correlation Coefficient (r)` = round(rho, 2),
  `df` = df,
  `p-value` = ifelse(p_value < 0.001, "< .001", round(p_value, 3))
)

# Print the correlation result in the console
print(correlation_result)

# Display the results in an APA-style table using knitr
kable(cor_table, 
      col.names = c("Variable 1", "Variable 2", "Correlation Coefficient (r)", "df", "p-value"), 
      caption = "Correlation Between Chicken Weight and Eggs Laid", 
      format = "pipe", 
      digits = 2)

```
```{r}
# Load necessary libraries
library(knitr)

# Prepare the data for the table
cor_table <- data.frame(
  Variable1 = "Chicken Weight",
  Variable2 = "Egg Size",
  `Spearman's rho` = round(rho, 2),
  `df` = df,
  `p-value` = ifelse(p_value < 0.001, "< .001", round(p_value, 3))
)

# Create the APA-style table using knitr
kable(cor_table, 
      col.names = c("Variable 1", "Variable 2", "Spearman's rho", "df", "p-value"), 
      caption = "Correlation Between Chicken Weight and Egg Size", 
      format = "pipe", 
      digits = 2)

```

```{r}

```

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Create a new column for average egg weight based on breed and egg size
chicken_data$Egg_Weight <- round(runif(nrow(chicken_data), 50, 80), 2)  # Random egg weights between 50 and 80 grams

# Calculate the average egg weight for each breed and egg size
egg_summary <- chicken_data %>%
  group_by(Breed, Egg_Size) %>%
  summarize(Average_Egg_Weight = mean(Egg_Weight), .groups = "drop")

# Create a bar plot with APA styling
ggplot(egg_summary, aes(x = Breed, y = Average_Egg_Weight, fill = Egg_Size)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +  # Adjust bar width for APA style
  scale_fill_manual(values = c("Small" = "#F4A300", "Medium" = "#0066CC", "Large" = "#33CC33", "X-Large" = "#FF6666")) +  # Custom colors
  labs(title = "Average Egg Weight by Chicken Breed and Egg Size", 
       x = "Chicken Breed", 
       y = "Average Egg Weight (grams)", 
       fill = "Egg Size") +
  theme_minimal(base_size = 12) +  # Set base font size
  theme(
    text = element_text(family = "Times New Roman"),  # Font style (Times New Roman for APA)
    axis.title = element_text(size = 12),  # Axis titles in font size 12
    axis.text = element_text(size = 10),  # Axis text in font size 10
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center and bold title
    legend.title = element_text(size = 12),  # Legend title font size
    legend.text = element_text(size = 10),  # Legend text font size
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for readability
    panel.grid.major = element_blank(),  # Remove major gridlines for cleaner look
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    panel.border = element_rect(color = "black", fill = NA)  # Add border around the plot for neatness
  )

```